---
title: "University of Washington Treatment Data Aggregation and Inspection"
author: "Micah Wright"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "../..")
```

# Purpose

This document saves the attribute tables from the UW treatment rasters as .csv files for later use.

# Setup

Load the necessary packages and functions.

```{r message=FALSE, warning=FALSE}
library(foreign)
library(data.table)
library(parallel)
```

# Load Data

Load treatment raster attribute tables. I exported these as .dbf files from within Arc, so no script exists. The column names were truncated during the export process, so they don't match the documentation exactly, but the column order is the same. 

First make a list of file names. Remove snags for now.

```{r}
file_list <- list.files("data/UW/batch_out",
                        pattern = ".dbf$",
                        full.names = TRUE)

no_treat <- c("data/UW/batch_out/Treatment_Snags.dbf")

file_list <- file_list[!file_list %in% no_treat]
```

Create a function to sum specified rows and convert to tons/acre.

```{r}
sum_residue <- function(dt, columns) {
        
        net_res <- Reduce("+", dt[, columns, with = FALSE]) 
        
        return(net_res/2000)
}
```

Write a function that loads the file, sums the appropriate columns, and saves the .csv.

```{r}
get_residue <- function(fullpath) {
        
        residue <- read.dbf(fullpath, as.is = TRUE)
        
        residue <- as.data.table(residue)
        
        treatment <- sub(".*_", "", fullpath)
        
        treatment <- sub("\\.dbf.*", "", treatment)
        
        # update 
        residue <- residue[, .(FCID2018 = Value,
                               TPA = TPA,
                               Stem_6t9_tonsAcre = sum_residue(dt = residue,
                                                               columns = c("CutStem6BL",
                                                                           "CutBarkSte")),
                               Stem_4t6_tonsAcre = sum_residue(dt = residue,
                                                               columns = c("CutStem4To",
                                                                           "CutBarkS_2")),
                               Stem_ge9_tonsAcre = sum_residue(dt = residue,
                                                                columns = c("CutStem6BG",
                                                                            "CutBarkS_1")),
                               Branch_tonsAcre = CutBranchB/2000,
                               Foliage_tonsAcre = CutFoliage/2000)]
        
        fwrite(residue, paste0("data/UW/residue/", treatment,".csv"))
}
```

Iterate over the file list after removing any old files.

```{r}
lapply(list.files(path = "data/UW/residue", full.names = TRUE), file.remove)

mclapply(file_list, 
         mc.cores = detectCores() - 1,
         function(x) get_residue(x))
```

