---
title: "FCCS"
author: "Micah Wright"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "../..")
```

# Purpose

This document pre-processes the FCCS data so it can be read in later.

### Dependencies

This document relies on several third-party packages, imported below.  

# Setup

Load the necessary packages and set ggplot2 theme.

```{r message=FALSE, warning=FALSE}
library(tidyverse) # loads core tidyverse packages
library(stringr) # for character string manipulation


theme_set(theme_classic() + 
                  theme(panel.grid.major = element_line(color = "grey90",
                                                        size = 0.2),
                        strip.background = element_blank()))
```

Load the FCCS fuel model data from Landfire. 

```{r}
FCCS <- read.csv("data/FCCS/tabular/LF_consume.csv", 
                 stringsAsFactors = FALSE,
                 header = FALSE)
```

The csv is in a Consume-specific format that we need to change for R to understand the column names. First extract the header rows

```{r}
KeepHead <- FCCS[1:2, ]
```

Now subset the rows so they only contain the loading values (no headers).

```{r}
FCCS <- FCCS[3:nrow(FCCS), ]
```

Ressign the column names, extracted above. Change the fuelbed number to numeric for later matching.

```{r}
names(FCCS) <- KeepHead[2, ]
```

Inspect the column names.

```{r}
names(FCCS)
```

The FCCS data has a lot of columns. Select the columns that we want to use and convert those columns to the appropriate class.

```{r}
FCCS <- dplyr::transmute(FCCS,
                  fuelbed_number = as.integer(fuelbed_number),
                  one_hr_sound = as.numeric(w_sound_0_quarter_loading),
                  ten_hr_sound = as.numeric(w_sound_quarter_1_loading),
                  hun_hr_sound = as.numeric(w_sound_1_3_loading),
                  oneK_hr_sound = as.numeric(w_sound_3_9_loading),
                  tenK_hr_sound = as.numeric(w_sound_9_20_loading), 
                  tnkp_hr_sound = as.numeric(w_sound_gt20_loading),
                  oneK_hr_rotten = as.numeric(w_rotten_3_9_loading), 
                  tenK_hr_rotten = as.numeric(w_rotten_9_20_loading),
                  tnkp_hr_rotten = as.numeric(w_rotten_gt20_loading),
                  litter_loading = as.numeric(litter_loading),
                  litter_depth = as.numeric(litter_depth),
                  duff_upper_loading = as.numeric(duff_upper_loading),
                  duff_upper_depth = as.numeric(duff_upper_depth),
                  duff_lower_loading = as.numeric(duff_lower_loading),
                  duff_lower_depth = as.numeric(duff_lower_depth),
                  lichen_depth = as.numeric(lichen_depth),
                  moss_depth = as.numeric(moss_depth))
```

Save it to a file.

```{r}
write_csv(FCCS, "data/FCCS/tabular/FCCS_fuelbed.csv")
```

