---
title: "Google Earth Engine Data Inspection"
author: "Micah Wright"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "../..")
```

Load the necessary packages.

```{r message=FALSE}
library(raster)
library(gdalUtils)
library(rgdal)
library(data.table)
```

Load the raw GEE data. It is 4km resolution fire weather data from Gridmet, summarized for different seasons for the years 2000-2018. There are state-level rasters for 1,000, 100, and 10-hour fuel moisture and 10m windspeed for each of three scenarios: 97th, 50th, and 37.5th percentile. The 37.5th percentile are for RX conditions, which were summarized for the months of September-October over the specified time period, all other rasters were summarized over June-September.

```{r}
rstack <- stack(list.files("data/GEE/raw", 
                           pattern = "*.tif", 
                           full.names = TRUE))
```

Plot the output, check to make sure spatial patterns seem reasonable and value distrbutions look right.

```{r}
plot(rstack, col = colorspace::diverge_hcl(n = 12, c = 100, l = c(50, 90), power = 1))
```

```{r}
hist(rstack)
```

Check a summary.

```{r}
knitr::kable(t(summary(rstack)))
```
