---
title: "SERC biomass scenario analysis"
author: "Micah Wright"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "../..")
```

# Purpose

This document loads the SERC scenario matrix and partitions it into lookup tables for later use.

# Setup

Load the necessary packages.

```{r message=FALSE}
library(readxl)
library(tidyverse)
```

# Extract Scenarios Lookup Table

Load the scenario matrix, and extract the different scenarios.

```{r message=FALSE}
scenarios <- read_csv("data/SERC/Scenario_Matrix_v11_model-input.csv",
                      col_names = FALSE)

scenarios_names <- as.character(scenarios[1,])

scenarios_names <- str_replace_all(scenarios_names, " ", "_")

scenarios <- scenarios[-1, 1:8]

names(scenarios) <- scenarios_names[1:8]

scenarios <- scenarios %>%
        filter(Slope == "LT40") %>%
        select(-Slope) %>%
        rename(Fraction_Piled = Fraction_Gross_Residues_That_Are_Piled,
               Fraction_Scattered = Fraction_Gross_Residues_That_Are_Scattered) %>%
        mutate(ID = as.integer(ID),
               Silvicultural_Treatment = gsub(" ", "_", 
                                              Silvicultural_Treatment),
               Fraction_Piled = gsub("[%]", "", Fraction_Piled),
               Fraction_Scattered = gsub("[%]", "", Fraction_Scattered),
               Burn_Type = ifelse(Burn_Type == "No", "None", Burn_Type)) %>%
        mutate(Silvicultural_Treatment = gsub("[%]", "", Silvicultural_Treatment)) 

noaction <- data.frame(ID = c(max(scenarios$ID) + 1,
                              max(scenarios$ID) + 2),
                       Silvicultural_Treatment = rep("No_Action", 2),
                       Fraction_Piled = rep("0", 2),
                       Fraction_Scattered = rep("0", 2),
                       Burn_Type = c("None", "Broadcast"),
                       Biomass_Collection = rep("None", 2),
                       Pulp_Market = rep("No", 2),
                       stringsAsFactors = FALSE)

scenarios <- bind_rows(scenarios, noaction)

write_csv(scenarios, "data/SERC/scenarios.csv")
```

# Extract Biomass Fate Lookup Tables

Creat a helper function that saves the scenario-specific biomass fate lookup tables after some minor munging to correct column names, etc.

```{r}
get_sections <- function(fraction_type, col_range) {
        
        section <- read_csv("data/SERC/Scenario_Matrix_v11_model-input.csv",
                      col_names = FALSE)
        
        section_names <- as.character(section[1,])
        
        section <- section[-1, c(1:8, col_range)]
        
        names(section) <- section_names[c(1:8, col_range)]
        
        section <- section %>%
                select(-contains("Bark"), -Stump, -Root) %>%
                transmute(ID = as.integer(ID),
                          Slope_Class = ifelse(Slope == "GT40", 80, 40),
                          Silvicultural_Treatment = gsub(" ", "_", `Silvicultural Treatment`),
                          Fraction_Piled = gsub("[%]", "",`Fraction Gross Residues That Are Piled`),
                          Fraction_Scattered = gsub("[%]", "", `Fraction Gross Residues That Are Scattered`),
                          Burn_Type = ifelse(`Burn Type` == "No", "None", `Burn Type`),
                          Biomass_Collection = `Biomass Collection`,
                          Pulp_Market = `Pulp Market`,
                          Stem_ge9 = as.numeric(gsub("[%]", "", `Stem 9+`))/100,
                          Stem_6t9 = as.numeric(gsub("[%]", "", `Stem 6-9`))/100,
                          Stem_4t6 = as.numeric(gsub("[%]", "", `Stem 4-6`))/100,
                          Branch = as.numeric(gsub("[%]", "", Branch))/100,
                          Foliage = as.numeric(gsub("[%]", "", Foliage))/100) %>%
                add_row(ID = rep(c(703, 704), 2),
                        Slope_Class = rep(c(40, 80), each = 2),
                        Silvicultural_Treatment = rep("No_Action", 4),
                        Fraction_Piled = rep("0", 4),
                        Fraction_Scattered = rep("0", 4),
                        Burn_Type = rep(c("None", "Broadcast"), 2),
                        Biomass_Collection = rep("None", 4),
                        Pulp_Market = rep("No", 4),
                        Stem_ge9 = rep(0.0, 4),
                        Stem_6t9 = rep(0.0, 4),
                        Stem_4t6 = rep(0.0, 4),
                        Branch = rep(0.0, 4),
                        Foliage = rep(0.0, 4)) %>%
                mutate(Silvicultural_Treatment = gsub("[%]", "", Silvicultural_Treatment))
        
        write_csv(section, paste0("data/SERC/lookup_tables/", fraction_type, ".csv"))

}
```

Create lists for fraction type and cell range.

```{r}
type_list <- list("recovered_by_size",
                  "harvested_biomass",
                  "piled_fuels",
                  "scattered_in_field")

range_list <- list(9:19,
                   20:30,
                   31:41,
                   42:52)
```

Iterate the function over the two lists.

```{r results=FALSE, message=FALSE}
map2(type_list, range_list, function(x, y) get_sections(x, y)) 
```

