---
title: "Tile size and cell count"
author: "Micah Wright"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "../..")
```

# Purpose

This document explores the tile size and cell count for each tile.

### Dependencies

This document relies on several third-party packages, imported below. 

# Setup

Load the necessary packages and set ggplot2 theme.

```{r message=FALSE, warning=FALSE}
library(raster) # raster data
library(rgdal) # raster and polygon 
library(parallel)
```

# Analysis

### Raster Import

FCID2018: UW FCID numbers. Can be reclassified to any UW output via attribute tables.

```{r}
FCID2018 <- raster("data/UW/FCID2018_masked.tif")
```

### Tile Import

Load the tiles. The size is customizable in make_tiles.R. Split the polygon into individual features, then remove the original.

```{r}
tiles <- readOGR("data/Tiles/clipped_tiles",
                 "clipped_tiles")

tile_list <- split(tiles, tiles$ID)

rm(tiles)
```

How big is each tile?

```{r}
area_fun <- function(shp) {
        cat("Area: ", area(shp)/10000, " Hectares")
}

area_fun(tile_list[[1]])
```

Make a function that crops the raster to each polygon and counts the cells that are not NA.

```{r}
get_count_fun <- function(ras, poly){
       rc <- crop(ras, poly)
       rows_ras <- nrow(as.data.frame(rc, na.rm = TRUE))
       return(rows_ras)
}
```

Apply the function to the list of raster file paths. This ouputs a list of rasters the size of the test zone, so all extents are the same. CRS and resolution were taken care of previously.

```{r}
rowcount_list <- mclapply(tile_list,
                      mc.cores = detectCores(),
                      function(i) get_count_fun(FCID2018, i))
```

Look at a histogram and summary of the values.

```{r}
rowcount <- unlist(rowcount_list)

hist(rowcount)

summary(rowcount)
```

